---
title: "Reproducible Research Assignment 1"
author: "S.B. Lim"
date: "Sunday, January 18, 2015"
output: html_document
---


##Loading and preprocessing the data

1.Load the data (i.e. read.csv())
``` {r echo=TRUE}
setwd("C:/Users/Soy Boon/Documents/R/Data/Reproducible Research/Assignment 1/repdata_data_activity")
data <- read.csv("activity.csv")
```

2.Process/transform the data (if necessary) into a format suitable for your analysis
``` {r echo=TRUE}
data$date <- as.Date(data$date, format = "%Y-%m-%d")
data$interval <- as.factor(data$interval)
```

##What is mean total number of steps taken per day?

1.Make a histogram of the total number of steps taken each day
``` {r echo=TRUE}
stepsPerDay <- aggregate(steps ~ date, data, sum)
colnames(stepsPerDay) <- c("date","steps")
hist(stepsPerDay$steps, main = paste("Total Number of Steps Taken Each Day"), xlab="Number of Steps Per Day")
```

2.Calculate and report the mean and median total number of steps taken per day
``` {r echo=TRUE}
stepsMean   <- mean(stepsPerDay$steps, na.rm=TRUE)
stepsMedian <- median(stepsPerDay$steps, na.rm=TRUE)
```

The Mean is `r stepsMean` and the Median is `r stepsMedian`

##What is the average daily activity pattern?

1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
``` {r echo=TRUE}
stepsPerInterval <- aggregate(steps ~ interval, data, mean)
plot(stepsPerInterval$interval,stepsPerInterval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
```

2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
``` {r echo=TRUE}
maxInterval <- stepsPerInterval[which.max(stepsPerInterval$steps),1]
```

The interval with the maximum number of steps is `r maxInterval`.

##Imputing missing values

1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
``` {r echo=TRUE}
sum(is.na(data))
```

2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Strategy <- Zeroes were used since it was the first day.

3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
``` {r echo=TRUE}
imputedData <- data
imputedData[as.character(imputedData$date) == "2012-10-01", 1] <- 0
```

4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r echo=TRUE}
imputedStepsPerDay <- aggregate(steps ~ date, imputedData, sum)
hist(imputedStepsPerDay$steps, main = paste("Total Number of Steps Taken Each Day"), xlab="Number of Steps Per Day")

imputedStepsMean <- mean(imputedStepsPerDay$steps, na.rm=TRUE)
imputedStepsMedian <- median(imputedStepsPerDay$steps, na.rm=TRUE)

imputedTotalDiff <- sum(imputedStepsPerDay$steps) - sum(stepsPerDay$steps)
```

The imputed Mean is `r imputedStepsMean` while the mean is `r stepsMean`.
The imputed Median is `r imputedStepsMedian` while the median is `r stepsMedian`.

There is no impact to the total daily number of steps with imputing the missing data.


##Are there differences in activity patterns between weekdays and weekends?

1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
``` {r echo=TRUE}
daytype <- function(date) {
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
        "weekend"
    } else {
        "weekday"
    }
}
data$daytype <- as.factor(sapply(data$date, daytype))
```

2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
``` {r echo=TRUE}
par(mfrow = c(2, 1))
for (type in c("weekend", "weekday")) {
    stepsType <- aggregate(steps ~ interval, data = data, subset = data$daytype == 
        type, FUN = mean)
    plot(stepsType, type = "l", main = type)
}
```



